# First code

Are you ready to write some code? I hope so, because it's time.

## Scripts and your workspace {.unnumbered}

In RStudio, click "File" in the upper left and then "New file -> R Script". This will open a window pane in the upper left of your screen. This "script" is basically a place for you to write and save code. Got it? Good.

Now click "File" then "Save as" and then save your script locally somewhere where you'll remember it. For example, you could save it on your "Desktop" as a file named "firstcode.R".

Your workspace consists of:

- Your script (upper left). This is where you talk to R.  
- The console (bottom left). This is where R talks back to you.  
- The "environment" (upper right). This is a place where you can explore data "objects" you've created (don't worry, we'll get there).  
- The "plots" tab (bottom right). This is where you'll see outputs of charts, maps, etc.  

## R as a calculator {.unnumbered}

At its most basic, R is a big calculator. Or a conversation partner. However you like to think of it. You talk to R by typing some code in your script (upper left), and R talks back by printing something in the console (bottom left).

For example, if you want to ask R what two plus two is, you'd write this on line one of your script.

```{r, eval = FALSE, echo = TRUE}
2 + 2
```

Then, if you place your cursor on that line and run "ctrl + enter" (or "cmd + enter" if on Mac), that code will "run" and R will give you the response.


```{r, eval = TRUE, echo = FALSE}
2 + 2
```

Cool, huh? You can, of course, type more complex math problems. Try typing the following on line 2:


```{r, eval = FALSE, echo = TRUE}
10392 * (13/72)^2 - sqrt(9.123) + (475/2.2)^3
```

Did you get the below as an answer?

```{r, eval = TRUE, echo = FALSE}
10392 * (13/72)^2 - sqrt(9.123) + (475/2.2)^3
```

Good.

## Objects {.unnumbered}


