# An early victory - interactive map with leaflet

#### Learning goals {-}

* Make a beautiful map with very few lines of code
* Feel great about it


### Let's make a map  {-}

Enough talk. Let's DO.

1. Open a script anywhere, name it anything, and save it.

2. On your first line of code, write `library(leaflet)`. What happened?

3. If necessary, in the CONSOLE, run `install.packages('leaflet')`.

4. Now re-run your first line of code.

5. You've just installed and then loaded the "leaflet" package. Don't know what that means? That's okay. Now install and load the "dplyr" package.

6. At this point you should have two lines of code. Now, add a third.

```{r,echo=TRUE,collapse=TRUE,message=FALSE,warning=FALSE, eval = FALSE}
leaflet()
```

Run all your code. What do you see?

7. Now add to your code as per the below.

```{r,echo=TRUE,collapse=TRUE,message=FALSE,warning=FALSE, eval = FALSE}
leaflet() %>%
  addTiles()
```

What do you see?

8. Now add some more.

```{r,echo=TRUE,collapse=TRUE,message=FALSE,warning=FALSE, eval = FALSE}
leaflet() %>%
  addTiles() %>% 
  setView(lng = 2.2,
          lat = 41.4, 
          zoom = 10)
```

9. Now change the latitude and latitude to your place of birth.

10. Go to https://leaflet-extras.github.io/leaflet-providers/preview/ and look at the pretty maps. Pick one that you like.

11. Replace `addtiles()` with `addProviderTiles('xxx')` where `xxx` is the name of the map style you like.
