# In-class survey

1. Open a new R script and name it `survey.R`

2. Write code to load the following libraries: `dplyr`, `readr`, and `ggplot2`

```{r, echo = FALSE, eval = TRUE}
library(dplyr)
library(readr)
library(ggplot2)
library(gsheet)
```

3. Fill out the survey at https://docs.google.com/forms/d/e/1FAIpQLSf_-POnYlMPAJO-RJxfbrZef6q9YAEQuPMb3qIBfcmizAkZAw/viewform?usp=sf_link.

4. Once everybody is done, read in the survey by running the below:

```{r, echo = FALSE, eval = TRUE}
if('survey.RData' %in% dir()){
  load('survey.RData')
} else {
survey <- gsheet::gsheet2tbl('https://docs.google.com/spreadsheets/d/1Ra5LhxdCRbmEp-wV07ZlREDt3gysBgyUG4AWuaa1pzk/edit?usp=sharing')
save(survey, file = 'survey.RData')
}
```


```{r, echo = TRUE, eval = FALSE}
survey <- gsheet::gsheet2tbl('https://docs.google.com/spreadsheets/d/1Ra5LhxdCRbmEp-wV07ZlREDt3gysBgyUG4AWuaa1pzk/edit?usp=sharing')
```


5. Run the below code to process the data a bit

```{r, echo = TRUE, eval = TRUE}
names(survey) <- c(
  'timestamp', #  "Timestamp", 
  'name', # "What is your name?", 
  'experience', # "How much experience do you have in R?", 
  'soda', # "How many sodas do you drink a week?", 
  'squat', # "How long do you think you can hold a wall squat (seconds)? (ie, the 'imaginary chair')?", 
  'egypt_or_france', # "Would you rather visit the pyramids of Egypt or the Eiffel tower of Paris?", 
  'bike', # "Do you know how to ride a bike?", 
  'swim', # "Do you know how to swim?", 
  'age', # "How old are you in years?", 
  'where', # "Where do you live?", 
  'lng', # "What is the longitude of your hometown?",
  'lat', # "What is the latitude of your hometown?", 
  'language', # "Do you know how to speak another language aside from English?", 
  'height', # "How tall are you in inches?", 
  'shoe', # "What size shoe do you wear? (US Size)", 
  'hand', # "Are you right-handed or left-handed?", 
  'happy', # "How happy do you consider yourself on a scale from 1 to 5, with 1 being extremely unhappy, 5 being extremely happy, and 3 being the average person?", 
  'sing', # "Do you consider yourself better or worse than average at singing?", 
  'pizza_or_love', # "What's more important pizza or love?", 
  'dance', # "Do you consider yourself better than average at dancing?", 
  'sleep', # "How many hours did you sleep last night", 
  'eat', # "Would you rather eat", 
  'adjective', # "What is the one adjective that best describes you?", 
  'teeth', # "How many teeth do you have?", 
  'braces', # "If you had to have braces the rest of your life, which color would you choose?", 
  'siblings', # "How many siblings do you have?", 
  'wrestle', # "Of the people in this room, who would win a pinky wrestle against Michael?", 
  'hangout', # "What is the best place to hang out in Sewanee?", 
  'movies', # "How many movies have you watched in the last month?"
  'dob' # "What is your date of birth?"
)

```

6. Run the following code to inspect the data

```{r, echo = TRUE, eval = FALSE}
View( survey )
```

## Exploring data {.unnumbered}

7. How many columns are in the data?

8. What are the names of the columns in the data?

9. How many rows are in the data?

10. Write code to look at just the first few rows of the data.

11. Write code to look at just the last few rows of the data.

12. How many people can ride a bike? Cannot ride a bike?

13. How many people can swim? Cannot swim?

14. How many teeth do people have?

15. How happy are people?

## R as a calculator {.unnumbered}

16. What is the average number of teeth per person?

17. What are the minimum and maximum number of teeth per person?

18. What are the average and median heights?

19. What percentage of people can ride a bike? Cannot ride a bike?

20. What percentage of people think that pizza is more important than love?

21. Summarize how much people slept last night.

22. Summarize how many sodas people drink per week.

## `dplyr` {.unnumbered}

Get in groups of two. Create 3 questions to ask of this survey data, each of which involves at least 2 variables. 

