# Exploring a dataset

Let's explore a dataset that consists of people who died and survived during a major event. Maybe, you can even guess the event.

1. Open a new R script and name it `survival.R`

2. Write code to load the following libraries: `dplyr`, `readr`, and `ggplot2`

```{r, echo = FALSE, eval = TRUE}
library(dplyr)
library(readr)
library(ggplot2)
```

3. Read in the data by running the following:

```{r, echo = TRUE, eval = TRUE}
people <- readr::read_csv('https://datatrain.cc/data/survival.csv')
```

4. How many columns are in the data?

5. What are the names of the columns in the data?

6. How many rows are in the data?

7. Write code to look at just the first few rows of the data.

8. Write code to look at just the last few rows of the data.

9. What do you suppose the `survived` variable refers to? 

10. Re-order the dataset by age, from youngest to oldest.

11. How old was the youngest person in the dataset?

12. Re-order the dataset by age, from oldest to youngest.

13. How old was the oldest person in the dataset?

14. How many males are there in the dataset?

15. How many females are there in the dataset?

16. What percentage of the dataset is female?

17. How many people survived?

18. How many people died?

19. What percentage of the people died?

20. How many males survived?

21. How many females died?

22. What percentage of males died?

23. What percentage of people who died were male?

24. What percentage of people who survived were male?

25. What percentage of people who died were female?

26. How many people were in each `class`?

27. How many people from each class survived and died?

28. What percentage of people from each class died?

29. What is the age of the youngest survivor?

30. What is the age of the oldest survivor?

31. What is the age of the youngest surviving male?

32. What is the sex of the oldest person who died?

33. What is the average of age of females who died?

34. What is the average age of males who survived?

35. Make a table showing the exact number of people that died and survived, grouping also by class and sex.

36. Create a variable named `outcome`. Populate this variable with the characters "survived" or "died" depending on the `survived` variable. 

37. Make another variable called `age_group`. If the person is over 18, call them "adult"; otherwise, call them "child".

38. Make another variable called `person_type`. Use `paste0` to combine the `sex` and `age_group` variables, in that order. 

39. Make another variable called `initial`. This should be the first letter of each person's last name. Hint, use `substr`.

40. What is the most prevalent `initial` in the dataset?

41. What is the least prevalent second letter in people's last names?

42. Use `nchar` to calculate the number of letters in each person's name, and make a variable with that value named `name_length`.

43. Who has the longest name in the dataset?

44. Who has the shortest name in the dataset?

45. Make a variable called `above_average_name_length`. This should be `TRUE` if the person's name length is above average, and false if below or at average.

46. Who is more likely to have survived: those with short names, or those with long names.

47. What is the age of the female with the shortest name who survived?

48. What is the sex of the oldest person to have died who had an above average name length?

49. Who, on average, has shorter names: males or females?

50. What is the total number of characters in all female names, and all male names?

51. Make a histogram of the age of people.

52. Make a barchart of the class of people.

53. Make a barchart of the `outcome` variable.

54. Make a scatterplot of the `name_length` and `age` variables.

55. Make a chart showing the age of males and females separately.

56. Make a chart showing survivorship by sex.

57. Make a chart showing survivorship by age.

58. Make a chart showing the age, sex, and survivor status of people.

59. Make a chart showing the distribution of the lengh of people's names.

60. Assume that the average life expectancy is 80. How many years of life were lost due to this event?

61. How many years of life lost were attributible to males? How many to females?

62. Make three more charts.

63. What caused this event?