# Exploring sanitation 

In the world of data science and public health, we have to work as a team to solve problems. Today you are going to apply what you've learned about `dplyr` and `ggplot` to learn about sanitation around the world. Good luck!

**1.** Make a folder in your datalab folder on your desktop and call it `sanitation`.

**2.** Open up an RScript and name it `sanitation.R` and save it into your `sanitation` folder.

**3.** In your RScript copy the following questions 4-7 and answer them as a team.

**4.** Look up improved vs. unimproved sanitation. What kind of sanitation does the US use, improved, unimproved or both?

**5.** What is open defecation?  Is it a type of improved or unimproved sanitation?

**6.** Is open defecation bad or good? Why or why not?

**7.** Why do some countries have high rates of open defecation and others do not?

**8.** Now that you know a bit about sanitation, let's start looking at some data. Download the dataset shared with you on slack (`sanitation.csv`) and save it to your `sanitation` folder.

**9.** **STOP! WE WILL DO THIS TOGETHER!** In your `sanitation.R` script underneath your questions, copy and paste the following line of code and inside the quotes hit tab. Then follow your files until you are in your `sanitation` folder which is inside your datalab folder on your desktop. If it is not there, go back and re-do questions 1 and 2 above (don't forget to comment your script!):

```{r, echo=TRUE, message=FALSE, eval = FALSE}
# set working directory
setwd("")
```

**10.** Write code to load the `tidyverse`.

**11.** Now load `sanitation.csv` into R and assign it a name using the following line of code:

```{r, echo=TRUE, message=FALSE, eval = FALSE}
# load in the data
sanitation <- read_csv('sanitation.csv')
```

**12.** What are the column names and based on these column names what is this dataset about? 

**13.** Make a table of the World Bank's income classifications.

**STOP! WE WILL DO THE NEXT FEW QUESTIONS TOGETHER!**

**14.** Since working with really long column names with spaces is really hard, let's change the column names together.

**15.** Let's look at data for the most recent year. 

**16.** To make it easier to explore the most recent data, let's name this dataset `latest`. 

**17.** Using `latest`, how many countries are there? 

**18.** How many countries are there from each continent?

**19.** How many countries are there from each income group?

**20.** Let's look at data for the low-income countries in Asia. How have their open defecation rates changed over time? **Let's talk through this together before writing any code!**

**RESUME WORKING WITH YOUR GROUP NOW!**

**21.** Make a line graph of safely managed sanitation services over the years in India.

**22.** Make a line graph of open defecation rates over the years for the country of your choice.

**23.** Using `latest`, make a scatterplot showing GDP per capita versus open defecation rates. Color the points by income group and give your plot good labels.

**24.** To make this scatterplot nicer, let's move the legend and reorder the income groups. We'll do this one together!

**25.** To learn another new thing, let's make a bar chart of sanitation services for South America. Then improve it by filling by income group.

**26.** Now make a bar chart of sanitation services for the continent of your choice. Remember to fill by income group!

