# Making some maps

Maps are great for exploring geographic data, like the prevalences of diseases in different countries or the
demographics of voting districts in different states. In addition to being efficient data visualizations, 
well made maps can be beautiful to look at!

1. Open a new R script and name it `maps.R`

2. Write code to load the following libraries: `dplyr`, `sf`, `spData`, `tmap`, and `tidycensus`

```{r, echo = FALSE, eval = TRUE}
library(dplyr)
library(sf)
library(spData)
library(tmap)
library(tidycensus)
```

3. Add the following code to your script and run it to make your first map:

```{r, echo = TRUE, eval = FALSE}
tm_shape( us_states ) + tm_polygons( )
```

4. What the heck is `us_states`? It's a geospatial dataset provided by the `spData` package, and you know from its name and the map above what it's for. To learn more, add the following code to your script and run it:

```{r echo = TRUE, eval = FALSE}
data( us_states )
View( us_states )
```

5. Let's make our first choropleth! Here's the code:

```{r echo = TRUE, eval = FALSE}
tm_shape( us_states ) + tm_polygons( col = "REGION")
```

6. Ack!! Whoever made this dataset forgot to check their spelling. Let's fix that and clean up our choropleth.

