# Cleaning Messy Data

### Learning goals {.unnumbered}

* This is a review exercise: learn the skills introduced in the previous modules by applying them to a universal data science scenario: cleaning up messy data.

## Your mission {.unnumbered}

Use the following code to load and peek at the following dataset: 

```{r echo=TRUE, eval = FALSE}
library(dplyr)
library(readr)
dives <- read_csv('https://raw.githubusercontent.com/databrew/intro-to-data-science/main/data/whales-dives.csv')

head(dives)
```

This dataframe is nice and tidy. Here are a few of its many tidy features:

* Each row is a single observation.

* There is not a single missing value anywhere in the dataset.

* The rows are organized from earliest to most recent, based on the data embedded in the `sit` column.

* Categorical columns have standardized formatting. In the `species` column, there are two levels: `HW` (humpback whale) and FW (fin whale). In the `behavior` column, there are also two levels: `FEED` and`OTHER`.

But this dataset was not always so pretty. Here is the **original** data file:

```{r echo=TRUE, eval = FALSE}
messy_dives <- read_csv('https://raw.githubusercontent.com/databrew/intro-to-data-science/main/data/whales-dives-messy.csv')

head(messy_dives)
```

**Your task in this review exercise is to write a script that carries out the necessary data cleaning steps to get this dataset from its original form to its tidy form.**

Test your work along the way, then demonstrate its completion, using the `identical()` function. If your `my_dives` version of the dataset is identical to the `dives` data above, the following logical test will be `TRUE`:

```{r echo=TRUE, eval=FALSE}
identical(dives,my_dives)
```

Enjoy!