# (PART) Review exercises {-}


#### Learning goals {-}

* This is a review exercise: apply the `dplyr` and `ggplot` skills introduced in the previous modules to explore a mystery dataset.  

&nbsp;  

```{r, results = 'asis', eval = ifelse(exists('is_teacher'),is_teacher,TRUE),echo=FALSE}
source('teacher_tip.R')

tip <- 'This review exercise is best done immediately after the module on **Dataframe wrangling**.'

teacher_tip(tip)
```

# A mystery dataset

Use the `dplyr` verbs to answer these questions.  

First, download the following dataset by running this code.  

```{r,echo=TRUE,collapse=TRUE, message = FALSE}
library(readr)
library(dplyr)
df <- read_csv('https://raw.githubusercontent.com/databrew/intro-to-data-science/main/data/deaths.csv')
```

**1.** Review the dataset and try to figure out which each row represents. "Each row is a ____."  

**2.** How many people are in the dataset?  

**3.** Use `summarize()` to count the number of men and women.  

**4.** Use `summarize()` to count the number of people in each class.   

**5.** Use `summarize()` to count the number of men and women in each class.   

**6.** What is the average age of men in the dataset?

**7.** What is the average age of women in the dataset? 

**8.** Use `mutate` to create a variable called `died`. This should be a boolean based on the `Survived` column (in which 1 means the person survived, and 0 means the person died).  

**9.** Use `mutate` to create a variable called `child`. This should be a boolean based on the `Age` column, indicating if someone was less than 18 years old.  

**10.** Create a different dataframe for men vs. women. Name them accordingly.  

**11.** Create a different dataframe for class 1, class 2, and class 3. Name them accordingly.  

**12.** For each of the 5 datasets you've just created, what is the death rate?  

**13.** For each of the 5 datasets, how many children died?  

**14.** Now, using the _full_ original dataset, calculate the child-specific death rate for each combination of class and sex (ie, "first class females", "third class males", etc.).  

**15.** What did you find? What might explain that?  

**16.** What is the average age of men and women, separately, in each class?

