---
title: Introduction to R
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: datatrainglobal/rbootcamp
description: "A resource & workbook for learning R"
favicon: img/favicon.ico
always_allow_html: true
---

```{r setup, include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               echo=FALSE,
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.width = 9.64,
               fig.height = 5.9,
               fig.path = 'figures/')
options(scipen=999)
```

# Welcome!

## What is this website? {.unnumbered}

This is not a textbook or an encyclopedia. This is not a reference manual. It is not exhaustive or comprehensive. We barely mention statistical tests at all. There is almost no theory. In fact, this curriculum is biased, non-representative, and incomplete -- *by design.*

**So what is this?** This website is a *guide* designed to help you through the most dangerous period in your data science education: your very first steps. The first few weeks. That is when 99% percent of people give up on learning to code.

You want to be part of the 1%? Okay, then you need to understand these 3 things:

**Premise 1: We learn best by doing.** Our goal is to get you *doing* data science. We will keep theory and detail to a minimum. We will give you the absolute basics, then offer you exercises and puzzles that motivate you to learn the rest. Then, once you've been *doing* data science for a bit, you soon begin *thinking* like a data scientist. By that, we mean tackling ambiguous problems with persistence, independence, and creative problem solving. With the focus on *doing* (as opposed to listening, reading, watching, etc.), the burden is on *you*. You gotta do the work.

**Premise 2: We learn best with purpose.** Once you gain comfort with the basic skills, you will be able to start working on real data, for real projects, with real impact. You will start to *care about what you are coding*. And that is when the learning curve *skyrockets* -- because you are motivated, and because you are learning *reactively*, instead of preemptively. Our goal is to get you to the point of take-off as quickly as possible.

**Premise 3: A simple toolbox is all you need to build a house.** Once you become comfortable with a few basic coding tools, you can build pretty much anything. The toolbox doesn't need to be that big; if you know how to use your tools well, and if you have enough building supplies (i.e., data), the possibilities are limitless.

**One more thing that this is not:** This is not a fancy interactive tutorial with bells or whistles. We purposefully designed this to be simple and "analog". You will not be typing your code into this website and getting feedback from a robot, or setting up an account to track your progress, or getting pretty merit badges or points when you complete each module.

Instead, you will be doing your work on your *own machine*, working with *real folders and files*, downloading data and moving it around, etc. -- all the things you will be doing as a real data scientist in the real world.

## Who this is for {.unnumbered}

This curriculum covers everything from the absolute basics of writing code in `R` to machine learning. As such, it is designed to be useful to everyone in some way. But the target audience for these tutorials is the *rookie*: the student who *wants* to work with data but has *zero* formal training in programming, computer science, or statistics.

## What you will learn {.unnumbered}

-   The **Getting started** unit teaches you how to use `R` (in `RStudio`). Here you will add the first and most important tools to your toolbox: working with variables, vectors, dataframes, scripts, and file directories.

-   More sections will go here...

 

```{r,include=TRUE,eval=FALSE}
library(dplyr)
library(Hmisc)
library(knitr)
files <- dir()
files <- files[grepl('.Rmd', files)]
files <- files[files != 'index.Rmd']
files <- gsub('.Rmd', '', files)
files <- Hmisc::capitalize(files)
files <- gsub('_', ' ', files)
files_split <- strsplit(files, split = '-')
chapter_numbers <- unlist(lapply(files_split, function(x){x[1]}))
chapter_names <- unlist(lapply(files_split, function(x){x[2]}))
toc <-
  dplyr::tibble(
    Number = chapter_numbers,
    Name = chapter_names
  )
toc <- toc[which(toc$Number != "Intro"),]
knitr::kable(toc)
```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

## Who are we? {.unnumbered}

Datatrain ([www.datatrain.global](https://datatrain.global)) is a project of Databrew ([www.databrew.cc](https://databrew.cc)).
